Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaOverview.tex
minor corrections and comments in Section 2.

Pilot-capable WMS enable high throughput of tasks execution while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. The design of PanDA WMS
started in 2005 to support ATLAS, one of the LHC experiments. PanDA went into
production for the LHC Run 1 on 2009, and was then extended with new subsystems
to be deployed on Run 2, on 2015.

@@

PanDA's usage model is based on multitenancy of resources and the support of at
least two types of HEP users: individual researchers and groups executing so
called `production' workflows. Users are free to submit tasks and workflows to
the PanDA WMS at any point in time, directly or via dedicated application
frameworks. Consistently, PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization (VO) abstraction.

@@

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}) but, at the
moment, they are not relevant to understand how PanDA has been ported to
supercomputers. For a full list of subsystems see Ref.~\cite{panda-wiki_url}.
Figure~\ref{fig:architecture} shows a diagrammatic representation of PanDA main
subsystems, highlighting the execution process of tasks while omitting
monitoring details to improve readability.

@@

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During the LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs
% with a fixed number of events
and then submitted to the PanDA Server.

@@

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
the application layer submits tasks descriptions to JEDI
(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
different size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA Server
for execution (Fig.~\ref{fig:architecture}:4).

@@

When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
module of the PanDA Server, requesting for a job to execute
(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
Buffer module for a job that is bound to the site of that pilot and ready to be
executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
upon availability, returns a suitable job the Job Dispatcher. The Job Dispatcher
dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).

>>>>

Pilot-capable WMS enable high throughput of tasks execution while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. The design of PanDA WMS
started in 2005 to support ATLAS. PanDA went into
production for the LHC Run 1 on 2009, and was then extended with new subsystems
for Run 2 in 2015.

@@

PanDA's usage model is based on multitenancy of resources and the support of at
least two types of HEP users: individual researchers and groups executing so
called `production' workflows. Users are free to submit tasks and workflows to
the PanDA WMS at any point in time, directly or via dedicated application
frameworks. PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization (VO) abstraction.

@@

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
ATLAS workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we
do not discuss them as they are  not relevant to  an understanding of how
PanDA has been ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Figure~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

@@

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs
% with a fixed number of events
and then submitted to the PanDA Server.

@@

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
the application layer submit task descriptions to JEDI
(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
different size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA Server
for execution (Fig.~\ref{fig:architecture}:4).

@@

When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
module of the PanDA Server, requesting  a job to execute
(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
Buffer module for a job that is bound to the site of that pilot and ready to be
executed. Task Buffer checks the global queue (i.e., the PanDA database) and,
upon availability, returns a suitable job the Job Dispatcher. \jhanote{end of previous sentence needs some attention} The Job Dispatcher
dispatches that job to the PanDA Pilot (Fig.~\ref{fig:architecture}:12).
